---
title: "ProjectNHL"
author: "Deepak Karawande"
date: "June 13, 2021"
always_allow_html: true
output:
  github_document:
    toc: yes
---

# Introduction
This vignette is an introduction on how to query REST-API end points using R and perform exploratory data analysis using various R packages. We’ll use be using the NHL REST-API endpoints which can be contracted by following instruction on [https://gitlab.com/dword4/nhlapi/-/blob/master/records-api.md](NHL Records API)

```{r include=FALSE}
knitr::opts_chunk$set(echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE)
```

# Packages
Following list of packages were used for accessing REST-API endpoints and exploratory data analysis and presentation.

```{r setup, echo=TRUE}
library("httr")
library("jsonlite")
library("tidyverse")
library("kableExtra")
```

# Accessing NHL Data

## Contract REST-API and create dataframe
NHLAPI project on github provides REST-API endpoints to access various datapoints for historical NHL games.

For this project, I accessed 7 different endpoints from NHLAPI to fetch information about NHL -
1. Franchise summary 
2. Franchise details 
3. Total stats for franchise 
4. Season records 
5. Skater records 
6. Admin history and retired numbers 
7. Team stats 

GET function from httr package was used for fetching data through REST-API. Using content and from JSON function data received was converted into r dataframe object.

```{r echo=TRUE}
# Base url to access NHLAPI 
get_baseURL <- function() { return("https://records.nhl.com/site/api") }

# helper function to fetch data from REST-API endpoint and conert it to dataframe.
fetch_DF <- function(api_url) {
  get_response <- GET(api_url)
  json_contents <- content(get_response, "text")
  list_res <- fromJSON(json_contents, flatten = TRUE)
  return(list_res)
}

# Get All Franchise
get_all_franchise <- function() {
  tab_name <- "franchise"
  full_url <- paste0(get_baseURL(), "/", tab_name)
  franchise_res <- fetch_DF(full_url)
  franchise_res$data
}
```

## Fetch data by Id or Name 
By default NHL REST-APIs return data for all franchise or team. Using helper functions I created a ability for user to provide Id or Name to fetch data for specific franchise/team if desired. If no Id or Name is passed query functions will return data for all franchise or teams.

```{r echo=TRUE}
# Function returns id for franchise by matching team common name to the fran_name argument provided. If team name was not found it will return -1.
get_franchise_id <- function(fran_id=NA, fran_name=NA) {
  retValue <- NA
  
  if(!is.na(fran_id)){
    return(fran_id)
  } else if(!is.na(fran_name)) {
    #get id from name
    franchise <- get_all_franchise() %>% filter(teamCommonName == fran_name)
    if(nrow(franchise) == 1) {
      retValue <- franchise$id
    } else {
      #we couldn't find the teamid for team name specified.
      retValue <- -1
    }
  }
  return(retValue)
}
```


```{r}


#Test Code
#get_franchise_id(fran_name='Eagles')
#get_franchise_id(fran_id=3)
#get_franchise_id()
#get_all_franchise()


get_baseURL_stats_teams <- function() {
  return("https://statsapi.web.nhl.com/api/v1/teams/")
}
# get All Teams

get_all_teams <- function() {
  full_url <- get_baseURL_stats_teams()
  franchise_res <- fetch_DF(full_url)
  franchise_res$teams
}

# Function returns id for team with common name matching to the argument provided. If team name not found it will return -1.
get_team_id <- function(team_id=NA, team_name=NA) {
  retValue <- NA
  
  if(!is.na(team_id)){
    return(team_id)
  } else if(!is.na(team_name)) {
    #get id from name
    team <- get_all_teams() %>% filter(teamName == team_name)
    if(nrow(team) == 1) {
      retValue <- team$id
    } else {
      #we couldn't find the teamid for team name specified.
      retValue <- -1
    }
  }
  return(retValue)
}

# Test Code
# get_team_id(team_id=2)
# get_team_id(team_name="Rangers")
```

## Function for fetching data of NHL records

With the help of functions above, created a R function to fetch data from desired NHL Records endpoint using either franchise id or team common name. If franchise id or team common name was not provided api will return data from all franchises.
Similar set of functions were used to contract with NHL team stats RES-API end point.

```{r echo=TRUE}
# Get NHL records as data.frame for given team by id/name or all. If team not found then empty data.frame is returned.
get_NHL_records <- function(tab_name, fran_id=NA, fran_name=NA) {

  full_url <- ""
  id_filter <- ""
  id <- get_franchise_id(fran_id, fran_name)
  
  if(!is.na(id)) {
    if(id != -1) {
    # fetch franchise information for given team id.
      if(tab_name == "franchise" | tab_name == "franchise-detail") {
        id_filter <- paste0("cayenneExp=id=",id)
      } else {
        id_filter <- paste0("cayenneExp=franchiseId=",id)
      }
      full_url <- paste0(get_baseURL(), "/", tab_name, "?", id_filter)
    } else {
      # Couldn't fine franchise by the name provided.
      return(data.frame())
    }
  } else {
    #fetch franchise information for all teams
    full_url <- paste0(get_baseURL(), "/", tab_name)
  }
  
  #fetch NHL records
  records_res <- fetch_DF(full_url)
  return(records_res$data)
}
```


```{r echo=FALSE, eval=TRUE}

#Test Code
# #get Franchise by id, name or all franchise if no name or id is provided.
# get_NHL_records("franchise")
# get_NHL_records("franchise", fran_id=17 )
# get_NHL_records("franchise", fran_name='Eagles' )
# 
# # Get Total Stats for given franchise by id/name or all.
# get_NHL_records("franchise-team-totals")
# get_NHL_records("franchise-team-totals", fran_id=17 )
# get_NHL_records("franchise-team-totals", fran_name='Eagles')
# 
# # Get seasonal records for Franchise by id/name or all.
# get_NHL_records("franchise-season-records")
# get_NHL_records("franchise-season-records", fran_id=17 )
# get_NHL_records("franchise-season-records", fran_name='Eagles')
# 
# # Get goalie records for Franchise by id/name or all.
# get_NHL_records("franchise-goalie-records")
# get_NHL_records("franchise-goalie-records", fran_id=17 )
# get_NHL_records("franchise-goalie-records", fran_name='Eagles')
# 
# # Get skater records for Franchise by id/name or all.
# get_NHL_records("franchise-skater-records")
# get_NHL_records("franchise-skater-records", fran_id=17 )
# get_NHL_records("franchise-skater-records", fran_name='Eagles')
# 
# # Get skater records for Franchise by id/name or all.
# get_NHL_records("franchise-detail")
#get_NHL_records("franchise-detail", fran_id=17 )
# get_NHL_records("franchise-detail", fran_name='Eagles')

```


```{r echo=FALSE, eval=TRUE}

# Get NHL stats as data.frame for given team by id/name or all. If team not found then empty data.frame is returned.
get_NHL_stats <- function(team_id=NA, team_name=NA) {
  full_url <- ""
  id <- get_team_id(team_id, team_name)
  
  if(!is.na(id)) {
    if(id != -1) {
      #fetch franchise information for given team id
      full_url <- paste0(get_baseURL_stats_teams(),id,"?expand=team.stats")
    } else {
      # no team found for this id return empty dataframe.
      return(data.frame())
    }
  } else {
    full_url <- get_baseURL_stats_teams()
  }
  list_total_stats <- fetch_DF(full_url)
  return(list_total_stats$teams)
}

#Test Code
#get_NHL_stats()
#get_NHL_stats(team_id=2)
#get_NHL_stats(team_name="Bruins")
#get_NHL_stats(team_name="Eagles")

```

## Wrapper function
Using Switch-Case, created wrapper function for providing simplicity for fetching data from all NHL REST-API endpoints relevant to this project, 
```{r echo=TRUE}

NHL_wrapper_api <- function(command, fran_id=NA, fran_name=NA, team_id=NA, team_name=NA) {
  result = switch(  
      command,  
      "get_franchise"= get_NHL_records("franchise", fran_id, fran_name),  
      "get_total_stats"= get_NHL_records("franchise-team-totals", fran_id, fran_name),  
      "get_season_records"= get_NHL_records("franchise-season-records", fran_id, fran_name),  
      "get_goalie_records"= get_NHL_records("franchise-goalie-records", fran_id, fran_name),
      "get_skater_records"= get_NHL_records("franchise-skater-records", fran_id, fran_name),
      "get_franchise_detail"= get_NHL_records("franchise-detail", fran_id, fran_name),
      "get_stats_for_team"=get_NHL_stats(team_id, team_name),
      print0("command not found, available commands are = \n",
             "get_franchise",
             "get_total_stats",
             "get_season_records",
             "get_goalie_records",
             "get_skater_records",
             "get_franchise_detail",
             "get_team_stats_for_season",
             "get_stats_for_team"
             )
  )
  return(result)
}
```


```{r echo=FALSE, eval=TRUE}

#Test Code
# NHL_wrapper_api(command="get_franchise")
# NHL_wrapper_api(command="get_total_stats")
# NHL_wrapper_api(command="get_season_records")
# NHL_wrapper_api(command="get_goalie_records")
# NHL_wrapper_api(command="get_skater_records")
# NHL_wrapper_api(command="get_franchise_detail")
# NHL_wrapper_api(command="get_stats_for_team")
# 
# NHL_wrapper_api(command="get_franchise", fran_id = 17)
# NHL_wrapper_api(command="get_total_stats", fran_id = 17)
# NHL_wrapper_api(command="get_season_records", fran_id = 17)
# NHL_wrapper_api(command="get_goalie_records", fran_id = 17)
# NHL_wrapper_api(command="get_skater_records", fran_id = 17)
# NHL_wrapper_api(command="get_franchise_detail", fran_id = 17)
# NHL_wrapper_api(command="get_stats_for_team", team_id = 2)
# 
# NHL_wrapper_api(command="get_franchise", fran_name = 'Eagles')
# NHL_wrapper_api(command="get_total_stats", fran_name = 'Eagles')
# NHL_wrapper_api(command="get_season_records", fran_name = 'Eagles')
# NHL_wrapper_api(command="get_goalie_records", fran_name = 'Eagles')
# NHL_wrapper_api(command="get_skater_records", fran_name = 'Eagles')
# NHL_wrapper_api(command="get_franchise_detail", fran_name = 'Eagles')
# NHL_wrapper_api(command="get_stats_for_team", team_name = 'Islanders')
```

# Exploratory Data Analysis

## Combine Franchise Summary and Details in single summary table.
To explore data, started with combining data from franchise summary and detail using inner_join to get basic tabular view of all franchises.
Rendered such summary table using kable function from knitr.

```{r echo=TRUE}
# Fetch franchise and franchise detail 
franchise <- as.tbl(NHL_wrapper_api(command="get_franchise"))
franchise_details <- as.tbl(NHL_wrapper_api(command="get_franchise_detail"))

# Combine franchise and franchise detail using inner join.
franchise_joined <- inner_join(franchise, franchise_details, by="id" ) %>%
  select(id, heroImageUrl, teamCommonName, active ) 

# make image urls renderable for franchise.
franchise_joined$heroImageUrl[!is.na(franchise_joined$heroImageUrl)] <- sprintf("![](%s){width=100px}", franchise_joined$heroImageUrl)

#print 
franchise_joined %>% 
  knitr::kable(caption="Franchise Summary Preview") %>% 
  kable_styling()

```

## Creating new variables
There are many ways to compute/add more variables to the dataset you are working with. I used group_by and summarize functions toc create two new variables totalWins and toalLosses for each unique combination of Franchise Id & Team.
Also computed percentage wins for each such combination of Franchise Id & Team using total wins and total losses. 

```{r echo=TRUE}
# calculate % of wins or % losses
Team_total_stats <- as.tbl(NHL_wrapper_api(command="get_total_stats"))

#colnames(Team_total_stats)
Team_total_stats %>% 
  select(teamName, franchiseId, wins, losses) %>% 
  group_by(franchiseId, teamName) %>%
  summarise(totalWins = sum(wins), totalLosses = sum(losses)) %>%
  mutate(perWins = round(totalWins/(totalWins+totalLosses),2)) %>%
  arrange(desc(perWins)) %>%
  head() %>%
  knitr::kable(caption="Win/Loss Percentages by Franchise & Teams") %>%
  kable_styling()
```
## Contingency tables
Contingency tables were created for Total goals scored by skaters by his position and franchise Id. 
2 separate tables were created by considering active and inactive players.

```{r echo=TRUE}
# create contingency table for franchise, Wins, losses, %wins
skaters <- as.tbl(NHL_wrapper_api(command="get_skater_records"))

skaters$positionCode <- factor(skaters$positionCode)
skaters$franchiseName <- factor(skaters$franchiseName)

levels(skaters$positionCode) <- c("Center Forward", "Defenseman", "Left Wing Forward", "Right Wing Forward")

skaters %>%
  filter(activePlayer == TRUE) %>%
  group_by(franchiseName, positionCode) %>%
  summarise(TotalGoals=sum(goals)) %>%
  spread(positionCode, TotalGoals) %>%
  head() %>%
  knitr::kable(caption="Active Players: Goal Counts by FranchiseID and Position") %>%
  kable_styling()

skaters %>%
  filter(activePlayer == FALSE) %>%
  group_by(franchiseName,positionCode) %>%
  summarise(TotalGoals=sum(goals)) %>%
  spread(positionCode, TotalGoals) %>%
  head() %>%
  knitr::kable(caption="Inactive Players: Goal Counts by FranchiseID and Position") %>%
  kable_styling()
```

## Numerical summeries
Numerical summaries were created for toals, gamesPlayed, mostGoalsOneGame, mostGoalsOneSeason by skaters with different positions as follows -
```{r echo=TRUE}
# Numerical Summaries

sakters_table <- function(pos){
  data <- skaters %>% filter(positionCode == pos) %>% select(goals, gamesPlayed, mostGoalsOneGame, mostGoalsOneSeason)
  kable(apply(data, 2, summary), caption = paste("Summary Goals by Position", pos), digit = 1) %>%
  kable_styling()
}

sakters_table("Center Forward")
sakters_table("Defenseman")
sakters_table("Left Wing Forward")
sakters_table("Right Wing Forward")

```

# Plots
ggplot packages supports creating nice plots to describe data.

## Bar plots
For created Bar plot of totals goals by skater's position used geom_bar function from ggplot with stat=identity to used y value provided for bar height.

```{r echo=TRUE}
skatersData <- skaters %>%
  group_by(positionCode) %>%
  summarise(TotalGoals=sum(goals), TotalGames=sum(gamesPlayed))

ggplot(skatersData, aes(x = positionCode, y=TotalGoals )) + 
  geom_bar(stat="identity") + 
  ggtitle("Bar Plot: Total Goals by PositionCode of Skaters")
```

## Histogram plots
Density plot is created using geom_histogram for mostSaves in one game by a goalie.

```{r echo=TRUE}
# fetch goalie data
goalie <- as.tbl(NHL_wrapper_api(command="get_goalie_records"))

#keep only mostSavesOneGame column in it
goalie_msg <- select(goalie, mostSavesOneGame, activePlayer)

goalie_msg$activePlayer <- factor(goalie_msg$activePlayer)
levels(goalie_msg$activePlayer) <- c("Inactive", "Active")
#remove NA rows
goalie_msg <- na.omit(goalie_msg)

ggplot(goalie_msg, aes(x = mostSavesOneGame, ..density..)) + 
  geom_histogram(bins = 20) + 
  ggtitle("Histogram for Most Save by a Goalie in one game") + 
  ylab("Density") + 
  geom_density(col = "red", lwd = 3, adjust = 0.4)
```

Using facet_wrap layer density plot of active and inactive player for mostSaves in one game by a goalies created as follows -

```{r echo=TRUE}
ggplot(goalie_msg, aes(x = mostSavesOneGame, ..density..)) + 
  geom_histogram(bins = 20) + 
  facet_wrap(~activePlayer) + 
  ggtitle("2 Histogram for Most Save by a Active vs InActive Goalie in one game") + 
  ylab("Density") + 
  geom_density(col = "red", lwd = 3, adjust = 0.4)

```

## Box plot
Box plot of points for active and inactive franchise is created with geom_boxplot layer.

```{r}

fStats <- NHL_wrapper_api(command="get_total_stats")
fStats$activeFranchise <- factor(fStats$activeFranchise)
levels(fStats$activeFranchise) <- c("Inactive", "Active")
ggplot(fStats, aes(x = activeFranchise, y = points)) + 
  geom_boxplot() + 
  geom_jitter(aes(color = activeFranchise)) + 
  ggtitle("Boxplot for Points")
```

## Scatter plot
geom_point layer function allows creating scatter plot with ggplot.
Her is active and inactive franchise wins and fit linear model line in it.
```{r}
# 
ggplot(fStats, aes(x = wins, y = points, group = activeFranchise)) + 
  geom_point(aes(color = activeFranchise)) + 
  geom_smooth(method = 'lm', color = 'green') + 
  ggtitle("wins vs points")

```

